---
title: "GBM(KLCI)"
author: "Lim Ming Feng"
date: "2026-02-06"
output: html_document
---

```{=html}

<style>
body {
  background-color: lightgrey; 

}
</style>
```

## Description

```{=html}

<p> I had download the some of indexes dataset from investing.com and saved it to my own project directory. I split my data into two sets one is training data, another is testing data based on the following date:

 <ol>training data (2014.1.1 - 2023.12.31)</ol>

 <ol>testing data (2024.1.1 - 2024.12.31)</ol>

</p>
```

## Data preprocessing

```{r}
library(lubridate)
library(xts)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### read dataset from csv

```{r}
data_name <- "FTSE China A50"
filepath<- paste0("C:\\Users\\User\\Programming\\R project\\dataset\\", data_name," Historical Data.csv")
data<- read.csv(filepath, header = TRUE ,stringsAsFactors = FALSE)

```

### convert date to proper format

```{r}
#2. convert date to proper format
data$Date <- as.Date(data$Date, format = "%m/%d/%Y") # if do not have this line the date cannot be visible 
data$Date <- ymd(data$Date)

```

### Clean numeric columns (remove columnns)

```{r}
data$Price <- as.numeric(gsub(",", "", data$Price))

```

## Data Understanding

```{r}
head(data) # first 5 data 
```

```{r}
str(data) # data structure
```

```{r}
head(data$Date)   # earliest dates
tail(data$Date)   # latest dates
any(duplicated(data$Date))   # FALSE is good
any(is.na(data$Date))        # should be FALSE
any(is.na(data$Price))       # if TRUE, decide how to handle (remove or impute)

```

### Drop unused data

```{r}
data_ <- data[, c("Date", "Price")] 

```

### Arrange data in chronological order

```{r}
data_ <- data_[order(data$Date), ] 
head(data_)
```

### Split data

```{r}
train_start <- ymd("2014-01-01")
train_end <- ymd("2023-12-31")
test_start <- ymd("2024-01-01")
test_end <- ymd("2024-12-31")

training_data <- data_[data_$Date >= train_start & data_$Date <= train_end, ]
testing_data <- data_[data_$Date >= test_start & data_$Date <= test_end, ]


```

## Parameter Estimation

### Compute Log Returns

simple return -\> mu = mu_log + (sigma\^2/2) log return -\> mu_log = mu-(sigma\^2/2)

in this case Compute simple return:

```         
(Price_Today/ Price_yesterday) - 1
```

```{r}
# simple return 
#data_sr <- (training_data$Price/lag(training_data$Price))-1

#sr <- na.omit(data_sr)
#options(digits =10)

#head(sr)

```

```{r}

# log return 
data_lr <- diff(log(training_data$Price))

options(digits =10)

head(data_lr)
```

```{r}
length(data_lr)

```

### Define variables

```{r}
S0 <- training_data$Price[nrow(training_data)] #use the last price from training data 
dt <- 1 #time increment 

# Assuming you have 'train_end' and 'test_end' defined as strings
# For test data 

num_of_wkdays <- tibble(
    date = seq(ymd(train_end) + days(1), ymd(test_end), by = "day")
  ) %>%
  mutate(
    is_weekday = wday(date, week_start = 1) <= 5 # Monday=1, Sunday=7. Check if day is <= Friday (5)
  ) %>%
  summarize(
    n_of_wkdays = sum(is_weekday)
  ) %>%
  pull()

num_of_wkdays
```

```{r}
T <- num_of_wkdays
N <- T/dt
t <- 1:N   # set up time step  
mu_log <- mean(data_lr) # daily mu 
sigma <- sd(data_lr)   # daily sigma 
num_of_sim <- 10000


# Generate random number
set.seed(2026)
b <- matrix(rnorm(N * num_of_sim, mean = 0, sd = 1), 
            nrow = N, 
            ncol = num_of_sim)


# cumulative sum function 
#W <- apply(b, MARGIN = 2, FUN = cumsum)


```

## Model Development and Simulation

```{r}
N_steps <- N +1  #include Day 0 (S0)
drift <- mu_log 

sim_prices <- matrix(NA, nrow = N_steps, ncol = num_of_sim)
sim_prices[1, ] <- S0 # start price S0 at first row

# Simulation loop
      for(i in 2:N_steps){
        diffusion <- (sigma*b[i-1, ])
        daily_log_return <- drift + diffusion
        sim_prices[i,] <- sim_prices[i-1, ] * exp(daily_log_return)    # t is equal to 1 
      }

head(sim_prices[, 1:5])


```

## Model Evaluation

no arbitrage condition

```{r}
# Calculate min , max prices per day across each simulation
S_max <- apply(sim_prices, MARGIN = 1, FUN = max)
S_min <- apply(sim_prices, MARGIN = 1, FUN = min)

# find midpoint 
S_pred <- 0.5*S_max + 0.5*S_min


# --- 1. Recreate the full list of all simulated weekdays ---

# Assuming 'train_end' and 'test_end' are defined as character strings 
# from your previous steps (e.g., "2023-12-31" and "2024-12-31")

all_sim_dates <- tibble(
  Date = seq(as.Date(train_end) + 1, as.Date(test_end), by = "day")
) %>%
mutate(
  Weekday_Name = weekdays(Date)
) %>%
filter(
  Weekday_Name %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
) %>%
select(Date)


# --- 2. Combine simulation results with these dates ---

# S_pred still needs the Day 0 value removed
sim_results_df <- tibble(
  Date = all_sim_dates$Date,
  pred = S_pred[-1] 
)





# --- 3. Perform an INNER JOIN to align real data and predictions by date ---

# This step keeps only the dates that exist in *both* data frames (skipping holidays)
final_df <- inner_join(
  testing_data %>% select(Date, real = Price), # Select real price and rename column
  sim_results_df,                              # Your simulation results
  by = "Date"
)

# calculate log returns for the real data 
final_df <- final_df %>%
  mutate(
    real_log_ret = log(real/lag(real))
  )

# Calculate Log Returns for the Predicted Prices

last_train_price <- tail(training_data$Price, 1)

final_df <- final_df %>%
  mutate(
    pred_log_ret = log(pred / lag(pred, default = last_train_price)) # mu_log + sigma*Z_i
  ) %>%
  na.omit() # Remove the first row if needed



# 3. Compute Metrics on Log Returns
mse_log <- mean((final_df$pred_log_ret - final_df$real_log_ret)^2)
rmse_log <- sqrt(mse_log)

# For MAPE, using log returns can be tricky (since they can be zero). 
# Usually, for returns, we just look at RMSE or MAE.
mae_log <- mean(abs(final_df$pred_log_ret - final_df$real_log_ret))

print(paste("Log-Return RMSE:", round(rmse_log, 6)))
print(paste("Log-Return MAE:", round(mae_log, 6)))

```

## Interpretation

```{r}

cat("On Average, GBM model predictions deviates about $", round(rmse_log,2), "from actual price.")

# can use rmse to divide S0 to check the relative error percentage 

cat("The has performed an accuracy of (",round(1 -mae_log,2)," %), the result has shown that only ", round(mae_log,2), "% error")
```

## Plotting

## GBM with MC simulation

```{r GBM and MC plot }
 # --- 1. Prepare Simulation Data for Plotting ---

# Add dates to the simulated prices matrix (from S0 start date)
all_sim_dates <- tibble(
  Date = seq(as.Date(train_end) + 1, as.Date(test_end), by = "day")
) %>%
  filter(weekdays(Date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>%
  add_row(Date = as.Date(train_end), .before = 1) # Add S0 date

# Convert the simulation matrix to a tidy (long) format DataFrame
sim_plot_data <- as.data.frame(sim_prices) %>%
  mutate(Date = all_sim_dates$Date) %>%
  pivot_longer(
    cols = -Date,
    names_to = "Scenario",
    values_to = "Price"
  )

# --- 2. Create the GBM MC Plot ---

ggplot() +
  # Plot all the individual simulation paths (light gray, thin lines)
  geom_line(data = sim_plot_data, 
            aes(x = Date, y = Price, group = Scenario, color = Scenario), 
            alpha = 0.5, size = 0.5 ) +
  
  # Plot the actual historical prices (blue, thick line)
  geom_line(data = final_df, 
            aes(x = Date, y = real), 
            color = "blue", size = 1.2) +
  
  # Plot your average prediction (red, thick dashed line)
  geom_line(data = final_df, 
            aes(x = Date, y = pred), 
            color = "red", size = 1, linetype = "dashed") +
  
  # Labels and theming
  labs(
    title = "Geometric Brownian Motion Monte Carlo Simulation",
    subtitle = paste("Forecast from", as.Date(train_end), "to", as.Date(test_end)),
    x = "Date",
    y = paste(data_name, "Index Price")
  ) +
  theme_minimal() + theme(legend.position = "none")
   
```

### Predicted data vs Actual data

```{r}
plot_data <- final_df %>%
  pivot_longer(
    cols = c(real_log_ret, pred_log_ret),
    names_to = "Type",
    values_to = "Price"
  ) %>%
  mutate(Type = recode(Type, 
                       "real_log_ret" = "Actual Log Return", 
                       "pred_log_ret" = "Predicted Log Return"))

# Then in ggplot:
ggplot(data = plot_data, aes(x = Date, y = Price, color = Type)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Actual Log Return" = "blue", "Predicted Log Return" = "red")) + 
  theme_minimal()+   
  labs(
    title = paste(data_name, "(Actual vs Predicted Log Returns)"),
    
  )
```

```{r}
# 1. Define your bounds based on your sigma and drift
# (Replace 'mu' and 'sigma' with your actual calculated values)
final_df <- final_df %>%
  mutate(
    lower_ci = drift - (1.96 * sigma),
    upper_ci = drift + (1.96 * sigma)
  )

# 2. Prepare the line data (keeping the CI columns separate)
plot_lines <- final_df %>%
  pivot_longer(
    cols = c(real_log_ret, pred_log_ret),
    names_to = "Type",
    values_to = "Price"
  ) %>%
  mutate(Type = recode(Type, 
                       "real_log_ret" = "Actual Log Return", 
                       "pred_log_ret" = "Predicted Log Return"))

# 3. Plot with geom_ribbon
ggplot() +
  # Add the shaded confidence interval first (so it's in the background)
  geom_ribbon(data = final_df, 
              aes(x = Date, ymin = lower_ci, ymax = upper_ci), 
              fill = "grey80", alpha = 0.5) +
  # Add the lines
  geom_line(data = plot_lines, aes(x = Date, y = Price, color = Type), linewidth = 0.8) +
  # Formatting
  scale_color_manual(values = c("Actual Log Return" = "blue", "Predicted Log Return" = "red")) +
  labs(title = "Log Returns with 95% Confidence Interval",
       subtitle = "Grey area represents the range where 95% of returns are expected to fall",
       y = "Log Return") +
  theme_minimal()

```

```{r}
# Reshape the data from wide format (real, pred columns) to long format (Type, Price columns)
# This is necessary for ggplot to easily draw two separate lines
plot_data <- final_df %>%
  pivot_longer(
    cols = c(real, pred), # Select the columns to pivot
    names_to = "Type",    # Name the new column 'Type' (will contain 'real' or 'pred')
    values_to = "Price"   # Name the value column 'Price'
  )

# Create the ggplot visualization
ggplot(data = plot_data, aes(x = Date, y = Price, color = Type)) +
  geom_line(size = 1) + # Draw the lines
  labs(
    title = "GBM Forecast vs. Actual Market Prices (Test Period = 2024)",
    x = "Date",
    y =  paste(data_name, "Index Price"),
    color = "Price Type"
  ) +
  scale_color_manual(values = c("real" = "blue", "pred" = "red")) + # Assign specific colors
  theme_minimal() # Use a clean theme

#stop("Stopping here")

```

## Residuals Plot

```{r}
library(patchwork)
## residuals plot 

# price residuals (heteroscedastic)
p1 <- ggplot(final_df, aes(x = pred, y = (real - pred))) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Price Residuals: The 'Fan'?", 
       x = "Predicted Price", y = "Error (unit index)") +
  theme_minimal()

# 2. Log Return Residuals (Likely Homoscedastic)
p2 <- ggplot(final_df, aes(x = pred_log_ret, y = (real_log_ret - pred_log_ret))) +
  geom_point(alpha = 0.5, color = "darkgreen") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Log Return Residuals: The 'Tube'?", 
       x = "Predicted Log Return", y = "Error (Log)") +
  theme_minimal()

p1 + p2

```
#  Hyperparameter tuning for drift
## rolling window 
```{r}
# n is the window size (e.g., last 60 trading days)
calc_static_mle_mu <- function(returns, n = 60) {
  # Simply calculate the average (MLE) of the last 'n' days
  static_mu <- mean(tail(returns, n), na.rm = TRUE)
  return(static_mu)
}

# Example: Get one single drift value from the end of your training data
mu_daily_tuned <- calc_static_mle_mu(data_lr, n = 60)
cat("Static MLE Mu: ", mu_daily_tuned)
```

## Simple Return


```{r}
# 1. Take the average of your log returns (The MLE)
avg_log_ret <- mean(data_lr, na.rm = TRUE)

# 2. Calculate the Volatility
sigma_val_ <- sd(data_lr, na.rm = TRUE)

# 3. TUNE THE DRIFT (Adding back the "Vol Drag")
mu_simple <- avg_log_ret + (0.5 * sigma_val_^2)
cat("Simple Return: " , mu_simple)
```

# Hyperparameter tuning for volatility


## Daily Realized Volatility Estimation -- (Static Rolling Windows == 60 days)

```{r}
# n is the window size (e.g., last 60 trading days)
calc_daily_realized_vol <- function(returns, n = 60) {
  # Simply calculate the standard deviation of the last 'n' days

  daily_rv <- sd(tail(returns, n))
  return(daily_rv)
}

# Example:
sigma_daily_tuned <- calc_daily_realized_vol(data_lr, n = 60)
cat("Daily Realized Volatility: ", sigma_daily_tuned)
```

## Daily EWMA Volatility

```{r}


calc_daily_ewma_vol <- function(returns, lambda = 0.94) {
  n <- length(returns)
  variance <- numeric(n)
  
  # Initialize with the historical variance of the returns
  variance[1] <- var(returns) 
  
  # EWMA Recursive Formula: σ²_t = λ * σ²_{t-1} + (1-λ) * r²_{t-1}
  for (i in 2:n) {
    variance[i] <- lambda * variance[i-1] + (1 - lambda) * (returns[i-1]^2)
  }
  
  # Take the square root of the latest variance
  latest_daily_vol <- sqrt(tail(variance, 1))
  return(latest_daily_vol)
}

# Example:
sigma_daily_ewma <- calc_daily_ewma_vol(data_lr)
cat("Daily EWMA Volatility: ", sigma_daily_ewma)

```

#MLE

```{r}
# 1. Load necessary libraries
if (!require("fitdistrplus")) install.packages("fitdistrplus")
library(fitdistrplus)

# 2. Prepare your data
# Replace 'your_price_vector' with your actual price data
actual_prices <- c(training_data$Price) 
train_log_ret <- diff(log(actual_prices))

# 3. Perform MLE (Fitting a Normal distribution to Log Returns)
# GBM assumes log returns follow a normal distribution N(mu, sigma^2)
mle_fit <- fitdist(train_log_ret, "norm", method = "mle")

# 4. Extract the Optimal Parameters
mu_mle <- mle_fit$estimate["mean"]
sigma_mle <- mle_fit$estimate["sd"]

# 5. daily parameters

mu_daily_mle <- mu_mle 
sigma_daily_mle <- sigma_mle

# Print results
cat("--- MLE Results ---\n")
print(mle_fit)
cat("\nDaily Drift (mu):", round(mu_daily_mle, 4))
cat("\nDaily Volatility (sigma):", round(sigma_daily_mle, 4))

# 6. Diagnostic Plot
plot(mle_fit)

```

##Simulation

```{r}
# --- Step 7: GBM Monte Carlo Simulation ---
set.seed(2026) # For reproducibility
n_paths <- 10000
n_days <- nrow(final_df)
S0 <- 1 # Normalizing starting price to 1 for easier error comparison

# Generate a matrix of random shocks (N x days)
Z <- matrix(rnorm(n_paths * n_days), ncol = n_days)

# Calculate simulated log-returns for each path
# Formula: (mu - 0.5 * sigma^2) * dt + sigma * sqrt(dt) * Z
# Correct: mu_mle already accounts for the -0.5*sigma^2 term


#tuning

# Drift

# mu_mle =mu_log 
# mu_simple 
# mu_daily_tuned (rolling 60 days)

# Volatility
# sigma_mle
# sigma_daily_tuned (rolling 60 days)
# sigma_daily_ewma

sim_log_rets <- mu_daily_tuned+ sigma_mle* Z

# Convert log-returns to cumulative price paths
# Use apply + cumsum to convert daily returns into price levels
sim_paths <- t(apply(sim_log_rets, 1, function(x) S0 * exp(cumsum(x))))

# Calculate daily min and max across ALL paths for each day
S_max_ <- apply(sim_paths, 2, max)
S_min_ <- apply(sim_paths, 2, min)

# Calculate your custom Midpoint prediction
pred_path <- 0.5 * S_max_ + 0.5 * S_min_

# Calculate the "Real" path from your actual log-returns
real_path <- S0 * exp(cumsum(final_df$real_log_ret))



```

## Evaluation

```{r}
# --- Step 8: Evaluation Metrics ---
# Root Mean Squared Error (RMSE)
rmse_val <- sqrt(mean((real_path - pred_path)^2))

# Mean Absolute Percentage Error (MAPE)
mape_val <- mean(abs((real_path - pred_path) / real_path)) * 100

cat("\n--- Evaluation Results ---")
cat("\nRMSE:", round(rmse_val, 6))
cat("\nMAPE:", round(mape_val, 2), "%\n")
```

## differences

```{r}

# --- Step 9: Visualization ---
plot(real_path, type = "l", col = "blue", lwd = 2, 
     ylim = range(c(real_path, sim_paths[1:10,])),
     main = "Actual vs. Predicted GBM Paths (2024)", 
     xlab = "Trading Days", ylab = paste(data_name, "Normalized Price"))

# Add a few individual simulation paths in light gray
for(i in 1:100) lines(sim_paths[i,], col = rgb(0.5, 0.5, 0.5, 0.2))

# Add the Predicted Mean path
lines(pred_path, col = "red", lwd = 2, lty = 1)

# Add Legend  
legend("topleft", legend = c("Real Data", "GBM Prediction", "Sim Paths"),
       col = c("blue", "red", "gray"), lty = c(1, 1, 1), lwd = 2)

```
